\rmv{Task Mining in Irregular Programs.}\new{Automatic Mining of Task Parallelism in Programs}

### 2017-02-22 docs/IEEE/paper.tex
Added Guido's modifications to the paper and the example of section 2

\guido{Not sure if we should dive into a discussion of what is a regular or an irregular program; to be even more precise we are actually mining task parallelism (for this moment) only within loops; hence I propose this new title}
\title{\rmv{Task Mining in Irregular Programs.}\new{Automatic Mining of Task Parallelism in Loops}}

@@

% CONTEXT
It is undeniable that parallel computing is nowadays the most pursued alternative by the developers  when striving for more application performance \cite{hwu2014ahead}. Two main reasons stand behind  the actual urge for parallel programming. \ed{First, the ever  increasing industry of multi-cores and many-cores architectures propitiates the development of software  with such behaviour \cite{haugen2016performance, vandierendonck2013analysis}.} \guido{Fine, but it seems to me a very convoluted way to say that this was due to the emergence of multicore/manycore  architectures as the alternative to maintain transistor power-density under control.} \rem{Second, programmers often code in simple parallel patterns  \cite{pinto2015large}, which evidences their awareness and will for parallelism}. \guido{Not sure if programming using patterns is an evidence that supports the urge for interest in parallel programming; on the contrary, it seems to me that it is more a consequence to tackle the complexity of the problem.}

@@

% PROBLEM
\ed{Programming parallel applications is a hard task, as programmers have to take care of many aspects of the job such as identify and keep variable dependences correct and assure synchronization and scalability of threads/processes while avoiding race conditions.} \ed{On the other hand, finding parallelism in \rem{sequentially-written} \guido{programs are typically "sequentially written"}  \ed{annotated sequential  code is not a simple task} as compiler static analyses is not precise enough to identify  parallel regions~\cite{kulkarni2007optimistic}. For example,} due  to the conservative trait of static analyses, loops are frequently marked as iteration-dependent when, in reality, \ed{their iterations might run in parallel most of the time. As a matter of fact,  many dependences pointed by the compiler might never occur during runtime, resulting in missing opportunities for parallelization.}

\new{A way to approach this problem is to use a programming model that can capture and enforce dependences between iterations at runtime when they occur. In such model  those iterations that do not have {\it loop-carried} dependencies are free to run in parallel, while iterations that dynamically create loop-carried dependences are serialized and dispatched in dependence order by  a system runtime.  In recent years, task-based execution models  have proven itself to be a scalable and quite flexible approach to extract regular and irregular parallelism from sequential code ~\cite{starss,bddt,cilk,sequoia,dandelion,legion,ooojava,dague}. In this model the programmer uses a {\it task} directive to mark  code regions in the program that are potential tasks and lists their corresponding dependecies.  OpenMP task scheduling~\cite{openmp2009}  is an example of such model. Although OpenMP task simplifies the mechanics of dispatching and running tasks, the burden of finding code regions to be  annotated with task directives still lies upon the programmer.}

@@

% SOLUTION
%Guido: This is an ambitious objective; ou work is more modest.
%Our work focuses on finding irregular code  and automatic generating compiler directives to make the parallelism explicit.
\ed{This paper presents an automatic technique that extracts task parallelism from loops which cannot be proved by static analysis to be loop-carried independent. } 
\rmv{It is a speculative technique that consists in the combination of static and dynamic analyses that increase 
the volume of tasks that can run in parallel.} \guido{Do not mention speculation here as it usually is applied to TLS or TM}

@@

>>>>

\guido{Not sure if we should dive into a discussion of what is a regular or an irregular program; to be even more precise we are actually mining task parallelism (for this moment) only within loops; hence I propose this new title}
\pedro{I think we should say "Automatic Mining of Task Parallelism in Programs" because we might mine task parallelism in recursive functions as well.}
\title{\rmv{Task Mining in Irregular Programs.}\new{Automatic Mining of Task Parallelism in Programs}}

@@

% CONTEXT
It is undeniable that parallel computing is nowadays the most pursued alternative by developers  when striving for more application performance \cite{hwu2014ahead}. Two main reasons stand behind  the actual urge for parallel programming. \ed{First, the ever  increasing industry of multi-cores and many-cores architectures propitiates the development of software  with such behaviour \cite{haugen2016performance, vandierendonck2013analysis}.} \guido{Fine, but it seems to me a very convoluted way to say that this was due to the emergence of multicore/manycore  architectures as the alternative to maintain transistor power-density under control.} \ed{Second, programmers often code in simple parallel patterns  \cite{pinto2015large}, which evidences their awareness and will for parallelism}. \guido{Not sure if programming using patterns is an evidence that supports the urge for interest in parallel programming; on the contrary, it seems to me that it is more a consequence to tackle the complexity of the problem.}

@@

% PROBLEM
Programming parallel applications is a hard task, as programmers ought to take care of many aspects of the job such as identifying and keeping variable dependences correct and assuring synchronization and scalability of threads/processes whilst avoiding race conditions. However, finding parallelism in programs that were not coded in a parallel paradigm \rmv{annotated sequential  code} is not a simple task, as compiler static analyses are not precise enough to identify  parallel regions \cite{kulkarni2007optimistic}. For example, due  to the conservative trait of static analyses, loops are frequently marked as iteration-dependent when, in reality, their iterations might run in parallel most of the time. As a matter of fact,  many dependences pointed by the compiler might never occur during runtime, resulting in several missing opportunities for parallelization.

A way to approach this problem is to use a programming model that can capture and enforce dependences between iterations at runtime whenever they occur. In such model, those iterations that do not have {\it loop-carried} dependencies are free to run in parallel, while iterations that dynamically create loop-carried dependences are serialized and dispatched in dependence order by  a system runtime.  In recent years, task-based execution models  have proven itself to be a scalable and quite flexible approach to extract regular and irregular parallelism from sequential code ~\cite{starss,bddt,cilk,sequoia,dandelion,legion,ooojava,dague}. In this model, the programmer uses a {\it task} directive to mark  code regions in the program that are potential tasks and lists their corresponding dependences.  The OpenMP task scheduling~\cite{openmp2009}  is an example of such model. Although OpenMP task simplifies the mechanics of dispatching and running tasks, the burden of finding code regions to be annotated with task directives still lies upon the programmer.

@@

% SOLUTION
%Guido: This is an ambitious objective; ou work is more modest.
%Our work focuses on finding irregular code  and automatic generating compiler directives to make the parallelism explicit.
\ed{This paper presents an automatic technique that extracts task parallelism from loops which cannot be proved by static analysis to be loop-carried independent. }

@@

%EXAMPLE - BFS
